---
title: "P_8105_HW3"
author: "Arthur Starodynov"
date: "2023-10-14-23"
output: github_document
---

Start the homework by loading the data set. 

```{r}
library(tidyverse)
library(p8105.datasets)
library(ggplot2)
library(scrutiny)
```



Here we start to examine the insta cart data sets that are available for use from the database for the class. 


```{r, }
insta_df <- data(instacart)
```

```{r}
instacart %>% 
  group_by(aisle) %>%
  summarize(
    n_obs = n()
    ) %>%
  arrange(desc(n_obs))
instacart %>%
  group_by(aisle_id)%>%
  summarize(n_obs = n()) %>%
  nrow()

instacart %>%
  group_by(aisle_id)%>%
  summarize(n_obs = n()) %>%
  max()
```

Through this R code above we can see that there are a total of 134 aisles with fresh vegetables being the most ordered from, followed by fresh fruits, and then packages vegetables fruits. 



```{r}
numbs_df = 
  instacart %>% 
  group_by(aisle) %>%
  summarize(
    n_obs = n()
    ) %>%
  filter(n_obs > 10000)

ggplot(numbs_df, aes(x = n_obs , y = aisle)) +
  geom_bar(fill = "purple", stat = "identity")

```


Above is a clean plot showing the number of observations seeen within each aisle when speicified each aisle to have more than 10,000 observations. 


```{r}
aisle_df = 
  instacart %>%
  group_by(aisle) %>%
  filter(aisle == c("baking ingredients" , "dog food care", "packaged vegetables fruits")) 

aisle_df %>% 
  group_by(add_to_cart_order) %>%
  summarize(count = n())


aisle_df

```

```

Below I have shown the three aisles shown with the most popular productrs bought from each

```


```{r} 
bake_df <- subset(instacart, aisle == "baking ingredients") %>% 
  group_by(product_name) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
head(bake_df)

dog_df <- subset(instacart, aisle == "dog food care") %>% 
  group_by(product_name) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
head(dog_df)

veggie_df <- subset(instacart, aisle == "packaged vegetables fruits") %>% 
  group_by(product_name) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
head(veggie_df)



```







## Problem 2


```
First, we do some data cleaning:
- format the data to use appropriate variable names
- focus on the “Overall Health” 
- include only responses from “Excellent” to “Poor”
- organize responses as a factor taking levels ordered from “Poor” to “Excellent”
```


```{r, warning = FALSE}
library(p8105.datasets)
data("brfss_smart2010")

brfss_df<- 
  brfss_smart2010 %>% 
  separate(Locationdesc, into = c("State", "County"), sep = "-") %>% 
  select(-c(Locationabbr)) %>%
  filter(Topic == "Overall Health", 
         Response %in% c("Poor", "Fair", "Good", "Very Good", "Excellent")) %>%
  mutate(Response = as.factor(Response))

```


```

Now with this information we can answer the question of In 2002, which states were observed at 7 or more locations? What about in 2010?

```

```{r} 
brfss_df %>%
  filter(Year == 2002) %>%
  group_by(State) %>%
  summarize(num = n_distinct(GeoLocation)) %>%
  filter(num >= 7)

brfss_df %>%
  filter(Year == 2010) %>%
  group_by(State) %>%
  summarize(num = n_distinct(GeoLocation)) %>%
  filter(num >= 7)

```


```

With this we can see which states in 2007 and 2010 were observed at 7 or more location. 


```


```{r, warning = FALSE, message = FALSE}
avg_df <- brfss_df %>%
    filter(Response == "Excellent") %>% 
    select(Year, State, GeoLocation, Data_value) %>%
    group_by(Year, State) %>%
    summarise(mean_data_value = mean(Data_value)) %>% 
    ungroup()
```


```

The dataset is shown only for excellent responses for the overall health questions, with the average across the locations.

```


```

Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).


```

```{r, warning = FALSE}
ggplot(avg_df, aes(x = Year, y = mean_data_value, color = State)) +
  geom_point(size = 1, alpha = .6) + 
  geom_line() +
  labs(y = "Avg Value", title = "Average Excellent Responses")

```



``` 

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```

```{r}
brfss_df %>% 
filter(Year %in% c(2006, 2010), 
         State == "NY ",
         !is.na(Response)) %>%
  select(Year, County, Response, Data_value) %>%
  arrange(Year, County, Response) %>%
  ggplot(aes(x = Response, y = Data_value)) +
  geom_boxplot() +
  facet_wrap(~Year)

```




## Problem 3



```
Load, tidy, merge, and otherwise organize the data sets. Your final dataset should include all originally observed variables; exclude participants less than 21 years of age, and those with missing demographic data; and encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).


```

```{r}

library(readr)

```


```{r, warning = FALSE, message = FALSE}
covar_df <- read_csv("./nhanes_covar.csv") %>%
  row_to_colnames(row = 4) %>%
  slice_tail(n = 250) %>%
  mutate(SEQN = as.integer(SEQN), 
         sex = as.factor(case_when(sex == 1 ~ "Male", sex == 2 ~ "Female")), 
         age = as.integer(age), 
         BMI = as.double(BMI), 
         education = as.factor(case_when(education == 1 ~ "Less than high school",
                                         education == 2 ~ "High school equivalent",
                                         education == 3 ~ "More than high school"))) %>%
  na.omit() %>%
  filter(age >= 21)

accel_df <- read_csv("./nhanes_accel.csv")

combined_df <- left_join(covar_df, accel_df, by = "SEQN")

```

```
Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items.

````

```{r}
combined_df %>% 
  group_by(education, sex) %>%
  summarize(`Number of People` = n()) %>%
  ungroup() %>%
  rename(Education = education, Sex = sex)

```


```

 The lowest education and sex combination is Females with a high school or high school-equivalent degre
 
```

```
Shown in graph below, females done with high school have higher median age than males and seen the same with the ones with same education as the males. 

```

```{r}
combined_df %>%
  select(SEQN, sex, age, education) %>%
  ggplot(aes(x = sex, y = age)) +
  geom_boxplot() +
  facet_wrap(~education)


```


```


Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.


```

```{r}
nhames_new_df<- combined_df %>%
  mutate(activity_per_day = rowSums(select(., contains("min")))) %>%
  select(SEQN:education, activity_per_day)

ggplot(nhames_new_df, aes(x = age, y = activity_per_day, color = sex)) +
  geom_point() +
  geom_smooth(se = F, method = "glm") +
  facet_wrap(~education) 

```





```
Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.
```


```{r}
combined_df %>% 
  pivot_longer(cols = contains("min"), names_to = "minute", values_to = "activity") %>%
  separate(col = minute, "n", into = c("min_char", "min")) %>%
  mutate(min = as.integer(min)) %>%
  ggplot(aes(x = min, y = activity, color = sex)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~education) 


```